---
title: "Mean ceramic dates for Caddo sites in the Neches River Basin"
author: "Robert Z. Selden, Jr."
date: "June 8, 2020"
output: github_document
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including all sites and vessels

The code chunk below was developed by Matthew A. Peeples, following Keith Kintigh, who was channelling Stanley South (1977).

```{r mcd, echo=TRUE}
# this code chunk was repurposed from Matthew A. Peeples excellent example (link above)
# set working directory
setwd(getwd())
mydata <- read.csv(file='nrb_mcd_data.csv',sep=',',header=T,row.names=1)
types <- as.matrix(colnames(mydata))

dates <- read.csv(file='nrb_type_dates.csv',sep=',',header=T)
bc <- min(dates[,2:3])
dates[,2:3] <- dates[,2:3]+(bc*-1)

midpt <- as.matrix(((dates[,3]-dates[,2])/2)+dates[,2])
dates <- cbind(dates,midpt)

mydata2 <- mydata


mcd.calc <- function(x,types,dates) {
tot <- as.matrix(rowSums(x))
for (i in 1:nrow(types)) {
for (j in 1:nrow(dates)) {
if (types[i,] == dates[j,1]) 
{x[,i] <- x[,i] * dates[j,4]}}}

mcd <- matrix(0,nrow(mydata),1)
rownames(mcd) <- rownames(mydata)

newtot <- as.matrix(rowSums(x))

for (i in 1:nrow(mcd)) {
mcd[i,1] <- newtot[i,]/tot[i,]+bc}
return(mcd)}

mcd <- mcd.calc(mydata2,types,dates)

###############################################################################

nsim <- 1000

data.rowsum <- as.matrix(rowSums(mydata))
range <- matrix(0,nrow(mydata),2)

for (i in 1:nrow(mydata)) {
data.sim <- rmultinom(nsim,data.rowsum[i,],prob=mydata[i,])
data.sim <- t(data.sim)
temp <- mcd.calc(data.sim,types,dates)
range[i,1] <- mean(temp) - (sd(temp)*1.96)
range[i,2] <- mean(temp) + (sd(temp)*1.96)}

output <- cbind(row.names(mydata),mcd,range)
colnames(output) <- c('site','mcd','lower','higher') #modified from Peeples' code
output

write.table(output,file='mcd_out.csv',sep=',',row.names=F)
# end of Peeples' code
```

## Plot results (mcd_out) with error bars

I prefer graphs over tables where possible (and appropriate), and I think it appropriate in this instance. 

```{r ggplot, out.width = "100%", dpi = 300, echo=TRUE, }
# plot mean ceramic dates with error bars
#install.packages("devtools")
#devtools::install_github("tidyverse/ggplot2")
library(tidyverse)
# read mcd_out
caddo<-read.csv("mcd_out.csv",header = TRUE, as.is=TRUE)
# define variables
Site<-caddo$site
Mean_Ceramic_Date<-caddo$mcd
lower<-caddo$lower
upper<-caddo$higher
# plot
cs = theme(
  axis.title.x = element_text(size = 14),
  axis.text.x = element_text(size = 8),
  axis.title.y = element_text(size = 14),
  axis.text.y = element_text(size = 6))

ggplot() +
  geom_errorbar(data=caddo, mapping=aes(x=Site, ymin=lower, ymax=upper), width=0.5, size=0.5, colour = "springgreen4") +
  geom_point(data=caddo, mapping=aes(x=Site, y=Mean_Ceramic_Date), size=2, shape=21, fill = "white") +
  coord_flip() + # provides a representation of time similar to what archaeologists are used to seeing
  cs
```

## Limit to those sites with five or more vessels

```{r mcd2, echo=TRUE}
# this code chunk was repurposed from Matthew A. Peeples excellent example (link above)
# set working directory
setwd(getwd())
mydata <- read.csv(file='nrb2_mcd_data.csv',sep=',',header=T,row.names=1)
types <- as.matrix(colnames(mydata))

dates <- read.csv(file='nrb2_type_dates.csv',sep=',',header=T)
bc <- min(dates[,2:3])
dates[,2:3] <- dates[,2:3]+(bc*-1)

midpt <- as.matrix(((dates[,3]-dates[,2])/2)+dates[,2])
dates <- cbind(dates,midpt)

mydata2 <- mydata


mcd.calc <- function(x,types,dates) {
tot <- as.matrix(rowSums(x))
for (i in 1:nrow(types)) {
for (j in 1:nrow(dates)) {
if (types[i,] == dates[j,1]) 
{x[,i] <- x[,i] * dates[j,4]}}}

mcd <- matrix(0,nrow(mydata),1)
rownames(mcd) <- rownames(mydata)

newtot <- as.matrix(rowSums(x))

for (i in 1:nrow(mcd)) {
mcd[i,1] <- newtot[i,]/tot[i,]+bc}
return(mcd)}

mcd <- mcd.calc(mydata2,types,dates)

###############################################################################

nsim <- 1000

data.rowsum <- as.matrix(rowSums(mydata))
range <- matrix(0,nrow(mydata),2)

for (i in 1:nrow(mydata)) {
data.sim <- rmultinom(nsim,data.rowsum[i,],prob=mydata[i,])
data.sim <- t(data.sim)
temp <- mcd.calc(data.sim,types,dates)
range[i,1] <- mean(temp) - (sd(temp)*1.96)
range[i,2] <- mean(temp) + (sd(temp)*1.96)}

output <- cbind(row.names(mydata),mcd,range)
colnames(output) <- c('site','mcd','lower','higher') #modified from Peeples' code
output

write.table(output,file='mcd_out2.csv',sep=',',row.names=F)
# end of Peeples' code
```

## Plot results (mcd_out) with error bars

I prefer graphs over tables where possible (and appropriate), and I think it appropriate in this instance. 

```{r ggplot2, out.width = "100%", dpi = 300, echo=TRUE, }
# plot mean ceramic dates with error bars
#install.packages("devtools")
#devtools::install_github("tidyverse/ggplot2")
library(tidyverse)
# read mcd_out
caddo2<-read.csv("mcd_out2.csv",header = TRUE, as.is=TRUE)
# define variables
Site<-caddo2$site
Mean_Ceramic_Date<-caddo2$mcd
lower<-caddo2$lower
upper<-caddo2$higher
# plot
cs = theme(
  axis.title.x = element_text(size = 14),
  axis.text.x = element_text(size = 8),
  axis.title.y = element_text(size = 14),
  axis.text.y = element_text(size = 6))

ggplot() +
  geom_errorbar(data=caddo2, mapping=aes(x=Site, ymin=lower, ymax=upper), width=0.5, size=0.5, colour = "springgreen4") +
  geom_point(data=caddo2, mapping=aes(x=Site, y=Mean_Ceramic_Date), size=2, shape=21, fill = "white") +
  coord_flip() + # provides a representation of time similar to what archaeologists are used to seeing
  cs
# end of code
```

